---
title: "Final Project"
author: "Capt Guadalupe Hernandez"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    keep_tex: true
geometry: margin=1in
fontsize: 10pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, results='hold', echo=FALSE, warning = FALSE, message=FALSE}
library(ggplot2)
library(car)
library(kableExtra)
library(knitr)
library(dplyr)
library(tidyr)
library(patchwork)
library(NSM3)
library(MASS)
library(cowplot)
library(reshape)
library(Sleuth3)
library(Hmisc)
library(truncnorm) 
options(digits = 10)
```

## Reproducing the Simulation
```{r, results='hold'}
compute_lambda_star = function(aA, bA, aB, bB, 
                               nA, nB, piA, piB, 
                               d, rho, Nsim = 5000) {
  
  # Generate observed trial data
  SA = rbinom(1, nA, piA)
  SB = rbinom(1, nB, piB)
  
  # Posterior sampling
  pA = rbeta(Nsim, aA + SA, bA + nA - SA)
  pB = rbeta(Nsim, aB + SB, bB + nB - SB)
  
  # Calc the diff of the posterior samples
  diff = pA - pB
  
  # Probability of Superiority (A > B + d)
  p_corr_star = mean(diff > d)
  
  # Probability of Ambiguity (-d <= A - B <= d)
  p_amb_star = mean(diff >= -d & diff <= d)
  
  # Composite decision index for Treatment A
  lambda_star_A = p_corr_star + (rho * p_amb_star)
  
  return(lambda_star_A)
}
```


```{r, results='hold'}
compute_xi = function(lambda_vals, theta = 0.9) {
  mean(lambda_vals > theta)
}
```


```{r, results='hold'}
run_scenario = function(scenario_name, piA, piB, 
                        aA_vec, bA_vec, aB_vec, bB_vec, 
                        n_vals = c(39, 65), d = 0.05, 
                        rho = 0.5, theta = 0.9, 
                        n_trials = 5000, Nsim = 5000) {
  
  results = data.frame()
  
  for (i in seq_along(aA_vec)) {
    # Store results for each sample size n_vals
    xi_results = sapply(n_vals, function(n) {
      # Generate N_trials of lambda_star
      lambda_samples = replicate(n_trials, {
        compute_lambda_star(aA_vec[i], bA_vec[i], aB_vec[i], bB_vec[i], 
                            nA = n, nB = n, piA = piA, piB = piB, 
                            d = d, rho = rho, Nsim = Nsim)
      })
      
      # Calculate xi
      return(compute_xi(lambda_samples, theta))
    })
    
    # Append to results data frame
    results = rbind(results, data.frame(scenario  = scenario_name,
                                        prior_set = i,
                                        prior_A   = paste0("(", aA_vec[i], ",", bA_vec[i], ")"),
                                        prior_B   = paste0("(", aB_vec[i], ",", bB_vec[i], ")"),
                                        xi_39     = 100 * xi_results[1],
                                        xi_65     = 100 * xi_results[2]))
  }
  
  return(results)
}
```


```{r, results='hold'}
# Set seed
set.seed(123)

scenario_1_1 = run_scenario(scenario_name = "1.1",
                            piA = 0.30, piB = 0.15,
                            aA_vec = 1,
                            bA_vec = 1,
                            aB_vec = 1,
                            bB_vec = 1,
                            d = 0.05,
                            rho = 0.5)
```


```{r, results='hold'}
# Set seed
set.seed(123)

scenario_1_2 = run_scenario(scenario_name = "1.2",
                            piA = 0.30, piB = 0.15,
                            aA_vec = c(3, 6, 9, 12, 15),
                            bA_vec = c(7, 14, 21, 28, 35),
                            aB_vec = c(2, 3, 5, 6, 8),
                            bB_vec = c(8, 17, 25, 34, 42),
                            d = 0.05,
                            rho = 0.5)
```


```{r, results='hold'}
# Set seed
set.seed(123)

scenario_1_3 = run_scenario(scenario_name = "1.3",
                            piA = 0.30, piB = 0.15,
                            aA_vec = c(3, 6, 9, 12, 15),
                            bA_vec = c(7, 14, 21, 28, 35),
                            aB_vec = c(2, 4, 6, 8, 10),
                            bB_vec = c(8, 16, 24, 32, 40),
                            d = 0.05,
                            rho = 0.5)
```


```{r, results='hold'}
# Set seed
set.seed(123)

scenario_1_4 = run_scenario(scenario_name = "1.4",
                            piA = 0.30, piB = 0.15,
                            aA_vec = c(3, 6, 9, 12, 15),
                            bA_vec = c(7, 14, 21, 28, 35),
                            aB_vec = c(1, 2, 3, 4, 5),
                            bB_vec = c(9, 18, 27, 36, 45),
                            d = 0.05,
                            rho = 0.5)
```


```{r, results='hold'}
# Set seed
set.seed(123)

scenario_1_5 = run_scenario(scenario_name = "1.5",
                            piA = 0.30, piB = 0.15,
                            aA_vec = c(5, 10, 15, 20, 25),
                            bA_vec = c(5, 10, 15, 20, 25),
                            aB_vec = c(4, 7, 11, 14, 18),
                            bB_vec = c(6, 13, 19, 26, 32),
                            d = 0.05,
                            rho = 0.5)
```


```{r, results='hold'}
# Set seed
set.seed(123)

all_results = rbind(scenario_1_1,
                    scenario_1_2,
                    scenario_1_3,
                    scenario_1_4,
                    scenario_1_5)

print(all_results)
write.csv(all_results, "sim_results.csv", row.names = FALSE)
```

## Proposed Extension

```{r}
# Updated simulation incorporating the Cost Ratio
run_scenario_extension = function(scenario_name, piA, piB, 
                                   aA_vec, bA_vec, aB_vec, bB_vec, 
                                   cost_ratios = c(4, 6, 9), # ratios
                                   n_vals = c(39, 65), d = 0.05, 
                                   rho = 0.5, n_trials = 5000, Nsim = 5000) {
  
  results = data.frame()
  
  for (i in seq_along(aA_vec)) {
    for (ratio in cost_ratios) {
      # Calculate formal Bayesian threshold
      theta = ratio / (1 + ratio)
      
      xi_results = sapply(n_vals, function(n) {
        lambda_samples = replicate(n_trials, {
          compute_lambda_star(aA_vec[i], bA_vec[i], aB_vec[i], bB_vec[i], 
                              nA = n, nB = n, piA = piA, piB = piB, 
                              d = d, rho = rho, Nsim = Nsim)
        })
        return(compute_xi(lambda_samples, theta))
      })
      
      results = rbind(results, data.frame(
        scenario  = scenario_name,
        cost_ratio = ratio,
        threshold = round(theta, 2),
        prior_A   = paste0("(", aA_vec[i], ",", bA_vec[i], ")"),
        prior_B   = paste0("(", aB_vec[i], ",", bB_vec[i], ")"),
        xi_39     = 100 * xi_results[1],
        xi_65     = 100 * xi_results[2]
      ))
    }
  }
  return(results)
}
```


```{r}
# Set the global seed
set.seed(123)

results_ext = run_scenario_extension(scenario_name = "Extended 1.1",
                                     piA = 0.30, piB = 0.15,
                                     aA_vec = c(1), bA_vec = c(1),
                                     aB_vec = c(1), bB_vec = c(1))
print(results_ext)
```


```{r}
results_ext2 = run_scenario_extension(scenario_name = "Extended 1.2",
                                     piA = 0.30, piB = 0.15,
                                     aA_vec = c(3, 6, 9, 12, 15), bA_vec = c(7, 14, 21, 28, 35),
                                     aB_vec = c(2, 3, 5, 6, 8),   bB_vec = c(8, 17, 25, 34, 42))

print(results_ext2)
```


```{r}
results_ext3 = run_scenario_extension(scenario_name = "Extended 1.3",
                                     piA = 0.30, piB = 0.15,
                                     aA_vec = c(3, 6, 9, 12, 15), bA_vec = c(7, 14, 21, 28, 35),
                                     aB_vec = c(2, 4, 6, 8, 10),   bB_vec = c(8, 16, 24, 32, 40))

print(results_ext3)
```


```{r}
results_ext4 = run_scenario_extension(scenario_name = "Extended 1.4",
                                     piA = 0.30, piB = 0.15,
                                     aA_vec = c(3, 6, 9, 12, 15), bA_vec = c(7, 14, 21, 28, 35),
                                     aB_vec = c(1, 2, 3, 4, 5),   bB_vec = c(9, 18, 27, 36, 45))

print(results_ext4)
```


```{r}
results_ext5 = run_scenario_extension(scenario_name = "Extended 1.5",
                                     piA = 0.30, piB = 0.15,
                                     aA_vec = c(5, 10, 15, 20, 25), bA_vec = c(5, 10, 15, 20, 25),
                                     aB_vec = c(4, 7, 11, 14, 18),   bB_vec = c(6, 13, 19, 26, 32))

print(results_ext5)
```


```{r, results='hold'}
ext_results = rbind(results_ext,
                    results_ext2,
                    results_ext3,
                    results_ext4,
                    results_ext5)

print(ext_results)
write.csv(ext_results, "ext_results.csv", row.names = FALSE)
```

## T Arm Treatment Extension

```{r}
# Function to get utilities for all arms in one trial
get_all_trial_utilities = function(pi_vec, n_vec, a_vec, b_vec, d, rho, Nsim) {
  # Generate outcomes for all arms
  S_vec = rbinom(length(pi_vec), n_vec, pi_vec)
  
  # Sample from all posteriors
  post_samples = matrix(NA, nrow = Nsim, ncol = length(pi_vec))
  for(i in 1:length(pi_vec)) {
    post_samples[, i] = rbeta(Nsim, a_vec[i] + S_vec[i], b_vec[i] + n_vec[i] - S_vec[i])
  }
  
  # Calculate all pairwise utilities (lambda_jk)
  # This returns a K x K matrix of lambda values
  K = length(pi_vec)
  lambda_matrix = matrix(0, K, K)
  
  for(j in 1:K) {
    for(k in 1:K) {
      if(j != k) {
        diff = post_samples[, j] - post_samples[, k]
        lambda_matrix[j, k] = mean(diff > d) + (rho * mean(abs(diff) <= d))
      }
    }
  }
  
  # Calculate U_j = average utility against all k != j
  U_vec = rowSums(lambda_matrix) / (K - 1)
  return(U_vec)
}
```


```{r}
select_treatment = function(U_vec) {
  which.max(U_vec)
}
```

### Runs

(Base-case) Weak prior

```{r}
# Parameters for a 3-arm trial
pi_vec = c(0.25, 0.30, 0.20)
n_vec  = c(65, 65, 65)
a_vec  = c(1, 1, 1)
b_vec  = c(1, 1, 1)
```


```{r}
set.seed(123)

# Run the simulation 10,000 times
trial_choices = replicate(10000, {
  # Get utilities for all arms
  U_vec = get_all_trial_utilities(pi_vec, n_vec, a_vec, b_vec, 
                                   d = 0.05, rho = 0.5, Nsim = 5000)
  
  # Select treatment
  select_treatment(U_vec)
})

# Summarize results
# Calculate probability of selecting each arm
table(trial_choices) / 10000
```


(Scenario A) Prior means correctly specified

```{r}
# Parameters for a 3-arm trial
pi_vec = c(0.25, 0.30, 0.20)
n_vec  = c(65, 65, 65)
a_vec  = c(1, 3, 2)
b_vec  = c(3, 7, 8)
```


```{r}
set.seed(123)

# Run the simulation 10,000 times
trial_choices = replicate(10000, {
  # Get utilities for all arms
  U_vec = get_all_trial_utilities(pi_vec, n_vec, a_vec, b_vec, 
                                   d = 0.05, rho = 0.5, Nsim = 5000)
  
  # Select treatment
  select_treatment(U_vec)
})

# Summarize results
# Calculate probability of selecting each arm
table(trial_choices) / 10000
```


(Scenario B) Moderately misspecified but highly informative prior specification

```{r}
# Parameters for a 3-arm trial
pi_vec = c(0.25, 0.30, 0.20)
n_vec  = c(65, 65, 65)
a_vec  = c(12, 3, 8)
b_vec  = c(28, 7, 32)
```


```{r}
set.seed(123)

# Run the simulation 10,000 times
trial_choices = replicate(10000, {
  # Get utilities for all arms
  U_vec = get_all_trial_utilities(pi_vec, n_vec, a_vec, b_vec, 
                                   d = 0.05, rho = 0.5, Nsim = 5000)
  
  # Select treatment
  select_treatment(U_vec)
})

# Summarize results
# Calculate probability of selecting each arm
table(trial_choices) / 10000
```

(Scenario C) Strongly informative and systematically misspecified prior specification

```{r}
# Parameters for a 3-arm trial
pi_vec = c(0.25, 0.30, 0.20)
n_vec  = c(65, 65, 65)
a_vec  = c(20, 3, 18)
b_vec  = c(20, 7, 32)
```


```{r}
set.seed(123)

# Run the simulation 10,000 times
trial_choices = replicate(10000, {
  # Get utilities for all arms
  U_vec = get_all_trial_utilities(pi_vec, n_vec, a_vec, b_vec, 
                                   d = 0.05, rho = 0.5, Nsim = 5000)
  
  # Select treatment
  select_treatment(U_vec)
})

# Summarize results
# Calculate probability of selecting each arm
table(trial_choices) / 10000
```


(D) Reversed / Imbalanced prior specification with strong anchoring on the best-performing arm

```{r}
# Parameters for a 3-arm trial
pi_vec = c(0.25, 0.30, 0.20)
n_vec  = c(65, 65, 65)
a_vec  = c(2, 15, 3)
b_vec  = c(8, 35, 17)
```


```{r}
set.seed(123)

# Run the simulation 10,000 times
trial_choices = replicate(10000, {
  # Get utilities for all arms
  U_vec = get_all_trial_utilities(pi_vec, n_vec, a_vec, b_vec, 
                                   d = 0.05, rho = 0.5, Nsim = 5000)
  
  # Select treatment
  select_treatment(U_vec)
})

# Summarize results
# Calculate probability of selecting each arm
table(trial_choices) / 10000
```



### 4 arm trials

(Scenario A) Prior means correctly specified

```{r}
# Parameters for a 4-arm trial
pi_vec = c(0.25, 0.30, 0.20, 0.1)
n_vec  = c(65, 65, 65, 65, 65)
a_vec  = c(1, 3, 2, 1)
b_vec  = c(3, 7, 8, 9)
```


```{r}
set.seed(123)

# Run the simulation 10,000 times
trial_choices = replicate(10000, {
  # Get utilities for all arms
  U_vec = get_all_trial_utilities(pi_vec, n_vec, a_vec, b_vec, 
                                   d = 0.05, rho = 0.5, Nsim = 5000)
  
  # Select treatment
  select_treatment(U_vec)
})

# Summarize results
# Calculate probability of selecting each arm
table(trial_choices) / 10000
```


### Stress test

```{r}
# Parameters for a 3-arm trial
pi_vec = c(0.15, 0.15, 0.15)
n_vec  = c(65, 65, 65)
a_vec  = c(3, 3, 3)
b_vec  = c(17, 17, 17)
```

```{r}
set.seed(123)

# Run the simulation 10,000 times
trial_choices = replicate(10000, {
  # Get utilities for all arms
  U_vec = get_all_trial_utilities(pi_vec, n_vec, a_vec, b_vec, 
                                   d = 0.05, rho = 0.5, Nsim = 5000)
  
  # Select treatment 
  select_treatment(U_vec)
})

# Summarize results
# Calculate probability of selecting each arm
table(trial_choices) / 10000
```


